---
title: "PML - Course Project"
author: "Diana Giraldo"
date: "08/23/2015"
output: html_document
---
Load the required packages:
```{r packs, message=FALSE}
library(dplyr)
library(caret)
library(ggplot2)
library(pander)
```

Load and convert the data:
```{r load_data, message=FALSE, warning=FALSE}
train <- read.table('pml-training.csv', header=TRUE, sep=",")
test <- read.table('pml-testing.csv', header=TRUE, sep=",")
train[, 12:159] <- lapply(train[, 12:159], function(x) as.numeric(as.character(x)))
test[, 12:159] <- lapply(test[, 12:159], function(x) as.numeric(as.character(x)))
```

### Pre Processing
Remove variables where the 95% of the values were NA's:
```{r rmnas}
rem <- which(colSums(is.na(train)) >= 0.95*nrow(train))
train <- train[,-rem]
test <- test[,-rem]
```

Remove variables with near to zero variance:
```{r nzv}
nzv <- nearZeroVar(train[,-60])
train2 <- train[,-nzv]
test2 <- test[,-nzv]
train2 <- train2[,-(1:6)]
test2 <- test2[,-(1:6)]
```

### Model Training
To estimate the out of sample error we set the training options to perform a K-Fold cross-validation with K=5:
```{r tropt}
trc <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
```

As this is a multi-class classification problem, we can use Decision Trees or Random Forests (RF), I chose RF because in most of the problems their performance is better than Trees.  
Due to the model takes a considerable amount of time to run, I save it in a file:
```{r modelfitsave, eval=FALSE, message=FALSE}
mod2 <- train(classe~., trControl = trc, data=train2 , method="rf")
save(mod2, file = "model_rf.rda")
```

### Predictions with the model
We load the file with the model:
```{r load_model}
load("model_rf.rda")
mod2
```

and predict the class for the testing set:
```{r predict, message=FALSE}
pred <- predict(mod2, test2)
pred
```

### Estimation of the out of sample error:
For each one of the 5-Fold cross validation we have an estimation of the out of sample error given by 1-Accuracy:
```{r err1, message=FALSE}
res <- mod2$resample
res <- mutate(res, err.est=1-Accuracy)
res
me.err <- mean(res$err.est)
sd.err <- sd(res$err.est)
```

#### The mean of the out-of-sample error is `r me.err` an the standard deviation is `r sd.err`